### 2024-06-06 journal/algorithm.tex
some more edits from final proofreading

Having defined the fields required for tasks \ref{findinroot}, \ref{findrank} and \ref{findenqueue},
we can easily see how to construct a new block $B$ during a \op{Refresh} in $O(1)$ \rev{steps}.
A \op{Refresh} on \node\ $\var{v}$ reads the values $h_{\ell}$ and~$h_{r}$ of the \fld{head} fields of $\var{v}$'s children and stores 
$h_{\ell}-1$ and $h_{r}-1$ in $B.\eleft$ and $B.\eright$.
Then, we can compute 
\begin{eqnarray*}
B.\fld{sum\sub{enq}}&=&\var{v}.\fld{left}.\fld{blocks}[B.\eleft].\fld{sum\sub{enq}} \\
&&+ \var{v}.\fld{right}.\fld{blocks}[B.\eright].\fld{sum\sub{enq}}.
\end{eqnarray*}
For a block $B$ in the root, $B.\fld{size}$ is computed using the \fld{size} field of the previous block $B'$ and
the number of \enqueues\ and \dequeues\ in~$B$
\rev{as follows.  Before the operations of $B$, the queue has $B'.\fld{size}$ elements.
The \enqueues\ of $B$ add $B.\fld{sum\sub{enq}}-B'.\fld{sum\sub{enq}}$ elements.
Then the \dequeues\ of $B$ remove $B.\fld{sum\sub{deq}} - B'.\fld{sum\sub{deq}}$ elements.
(If subtracting this many elements would make the new size negative because there are more \dequeues\ than there are elements, then the new size is 0.)  Thus, we have}
\begin{eqnarray*}
B.\fld{size} &=& \max(0, B'.\fld{size} + (B.\fld{sum\sub{enq}}-B'.\fld{sum\sub{enq}})\\ 
			&& \hspace*{19mm}\mbox{ } - (B.\fld{sum\sub{deq}} - B'.\fld{sum\sub{deq}})).
\end{eqnarray*}

The only remaining field is $B.\fld{super}$.  When the block 
$B$ is created for a \node\ $\var{v}$, we do not yet know where its
superblock will eventually be installed in \var{v}'s parent.
So, we leave $B.\fld{super}$ blank.  
Soon after $B$ is installed,
%a call to \op{Advance} at line \ref{helpAdvance} or \ref{advance} by 
some process will
set \var{B}.\fld{super} to a value read from the \fld{head} field of \var{v}'s parent.
We shall show that this happens soon enough that $B.\fld{super}$ can differ from the true index of \rev{$B$'s superblock}
by at most 1.
%After $B$ is installed 
%in $\var{v}.\fld{blocks}[\var{h}]$, processes cooperate to fill in $B.\fld{super}$ 
%when they attempt to advance $\var{v}.\fld{head}$ from $h$ to $h+1$,
%using a value they read from the \fld{head} field of $\var{v}$'s parent.
%As mentioned above, this might not be the exact index of $B$'s superblock, but we
%shall prove that it is close.

@@

\Function{void}{Enqueue}{\typ{Object} \var{e}} 
    \State \hangbox{let \var{B} be a new \typ{Block} with fields \fld{element} \assign\ \var{e},
		\fld{sum\sub{enq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{enq}}+1$,\\
		\fld{sum\sub{deq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{deq}}$}\label{enqNew}
    \State \Call{Append}{\var{B}}\Comment{\rev{append \op{Enqueue} to \var{leaf}.\fld{blocks} and propagate to root}}\label{enqAppend}
\EndFunction{Enqueue}

@@

\Function{Object}{Dequeue}{} 
    \State \hangbox{let \var{B} be a new \typ{Block} with fields \fld{element} \assign\ \nl,
	    \fld{sum\sub{enq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{enq}}$,\\
	    \fld{sum\sub{deq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{deq}}+1$}\label{deqNew}
    \State \Call{Append}{\var{B}}\Comment{\rev{append \dequeue\ to \var{leaf}.\fld{blocks} and propagate to root}}\label{deqAppend}
    \State $\langle \var{b}, \var{i}\rangle$ \assign\ \Call{IndexDequeue}{\var{leaf}, $\var{leaf.}\head-1$, $1$} \Comment{\rev{find where \op{Deqeueue} is in \var{root}.\fld{blocks}}}\label{invokeIndexDequeue}
    \State \Return{\Call{FindResponse}{\var{b, i}}}\Comment{\rev{find response \dequeue\ should return}}\label{deqRest}
\EndFunction{Dequeue}

@@

\Function{void}{Propagate}{\typ{Node} \var{v}} \Comment{propagate blocks from \var{v}'s children to root}
    \If{{\keywordfont not} \Call{Refresh}{\var{v}}} \label{firstRefresh}  \hfill \Comment{double refresh:  \rev{if first fails, try again}}
        \State \Call{Refresh}{\var{v}} \label{secondRefresh}
    \EndIf
    \If{$\var{v} \neq \var{root}$} \hfill \Comment{recurse up tree}
        \State \Call{Propagate}{\var{v}.\fld{parent}}
    \EndIf
\EndFunction{Propagate}

@@

\begin{figure*}
\begin{algorithmic}[1]
\setcounter{ALG@line}{39}
\Function{Block}{CreateBlock}{\typ{Node} \var{v}, \typ{int} \var{i}} \Comment{create new block $B$ for \op{Refresh} to install in \var{v}.\fld{blocks}[\var{i}]}
    \State \hangbox{let \var{B} be a new \typ{Block} with fields\label{initNewBlock}
    	\eleft\ \assign\ $\var{v}.\fld{left}.\head - 1$,\label{createEndLeft}
    	\eright\ \assign\ $\var{v}.\fld{right}.\head - 1$,\label{createEndRight}\\
		\fld{sum\sub{enq}} \assign\ \var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{enq}} + \var{v}.\fld{right.blocks}[\var{B}.\eright].\fld{sum\sub{enq}}\label{createSumEnq},\\
		\fld{sum\sub{deq}} \assign\ \var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{deq}} + \var{v}.\fld{right.blocks}[\var{B}.\eright].\fld{sum\sub{deq}}}\label{createSumDeq}
    \State \var{num\sub{enq}} \assign\ $\var{B}.\fld{sum\sub{enq}} - \var{v}.\fld{blocks}[\var{i}-1].\fld{sum\sub{enq}}$\label{computeNumEnq}
    \State \var{num\sub{deq}} \assign\ $\var{B}.\fld{sum\sub{deq}} - \var{v}.\fld{blocks}[\var{i}-1].\fld{sum\sub{deq}}$
    \If{$\var{num\sub{enq}} + \var{num\sub{deq}} = 0$} \Return \nl\ \Comment{no blocks need to be propagated to \var{v}}\label{testEmpty} 
    \ElsIf{$\var{v} = \var{root}$} \var{B}.\fld{size} \assign\ max(0, $\var{v}.\fld{blocks}[\var{i}-1].\size\ + \var{num\sub{enq}} - \var{num\sub{deq}}$)\label{computeLength}
    \EndIf
    \State \Return \var{B}
\EndFunction{CreateBlock}

@@

    \If{\var{v} is a leaf node} \Return \var{v}.\fld{blocks}[\var{b}].\fld{element} \label{getBaseCase}
    \Else 
        \State \var{sum\sub{left}} \assign\ \var{v}.\fld{left.blocks}[\var{v}.\fld{blocks}[\var{b}].\eleft].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}$] from \var{v}.\fld{left}}
        \State \var{prev\sub{left}} \assign\ \var{v}.\fld{left.blocks}[\var{v}.\fld{blocks}[$\var{b}-1$].\eleft].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}-1$] from \var{v}.\fld{left}}
        \State \var{prev\sub{right}} \assign\ \var{v}.\fld{right.blocks}[\var{v}.\fld{blocks}[$\var{b}-1$].\eright].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}-1$] from \var{v}.\fld{right}}
        \If{$\var{i} \leq \var{sum\sub{left}} - \var{prev\sub{left}}$} \fld{dir} \assign\ \fld{left}\label{leftOrRight} \Comment{required \enqueue\ is in \var{v}.\fld{left}}
        \Else \Comment{required \enqueue is in \var{v}.\fld{right}}
            \State \fld{dir} \assign\ \fld{right}
            \State $\var{i}\ \assign\ \var{i} - (\var{sum\sub{left}} - \var{prev\sub{left}})$ \Comment{\rev{deduct \# \enqueues\ in $v.\fld{blocks}[b]$ that came from $v.\fld{left}$}}
        \EndIf \label{endChooseDir}
        \State \linecomment{find \enqueue\'s block in \var{v}.\fld{dir.blocks} and its rank within block}
        \State \hangbox{use binary search to find minimum $\var{b}'$ in range [\var{v}.\fld{blocks}[$\var{b}-1$].\edir+1..\var{v}.\fld{blocks}[\var{b}].\edir] such\\
        	 that $\var{v}.\fld{dir.blocks}[\var{b}'].\fld{sum\sub{enq}} \geq \var{i} + \var{prev\sub{dir}}$\label{getChild}}
        \State $\var{i}'$ \assign\ $\var{i} - (\var{v}.\fld{dir.blocks}[\var{b}'-1].\fld{sum\sub{enq}} - \var{prev\sub{dir}})$\label{getChildIndex}
        \State \Return\Call{GetEnqueue}{\var{v}.\fld{dir}, $\var{b}'$, $\var{i}'$} \label{getRecurse}
    \EndIf
\EndFunction{GetEnqueue}

@@

\opemph{CreateBlock}(\var{v, i}) is used
by \op{Refresh} to construct a new block $B$ to be installed in \var{v}.\fld{blocks}[\var{i}].
The \eleft\ and \eright\ fields store the indices of the last blocks appended to \var{v}'s
children, obtained by reading the \fld{head} index in \var{v}'s children.
Since the \fld{sum\sub{enq}} field should store the number of \enqueues\ in
\var{v}.\fld{blocks}[1..\var{i}] and these \enqueues\ come from \var{v}.\fld{left.blocks}[1..\var{B}.\eleft] and \var{v}.\fld{blocks}[1..\var{B}.\eright], line \ref{createSumEnq} sets
\fld{sum\sub{enq}} to the sum of $\var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{enq}}$ and $\var{v}.\fld{right}.\fld{blocks}[\var{B}.\eright].\fld{sum\sub{enq}}$.
Line \ref{computeNumEnq} sets \var{num\sub{enq}} to the number of \enqueues\ in the new block by
subtracting  the number of \enqueues\ in \var{v}.\fld{blocks}[$1..\var{i}-1$] from \var{B}.\fld{sum\sub{enq}}.
The values of \var{B}.\fld{sum\sub{deq}} and \var{num\sub{deq}} are computed similarly.
\rev{If the new block contains no operations, \op{CreateBlock} returns \nl\ to indicate
 there is no need to install~it.
Otherwise,} if \var{B}
is going to be installed in the root, line \ref{computeLength} computes the \fld{size} field, which
represents the number of elements in the queue after the operations in the block are performed.

@@

To compute the response of the $i$th \dequeue\ in the $b$th block
of the root, \opemph{FindResponse}(\var{b, i}) determines at line \ref{checkEmpty} if the queue is empty.
If not, line \ref{computeE} computes the rank \var{e} of the
\op{Enqueue} whose argument is the \dequeue's response. 
\rev{The routine next searches for}
the index $b_e$ of the block in \var{root}.\fld{blocks} that contains 
the \var{e}th \enqueue.
Since the \enqueue\ is linearized before the \dequeue, $b_e\leq b$.  To find the left end of the range \rev{to search for $b_e$, \op{FindResponse} does} a doubling search \cite{BY76}, comparing \var{e} to the \fld{sum\sub{enq}} fields at indices $b-1, b-2, b-4, b-8, \ldots$.
\rev{Then, the doubling search performs a binary search on the \fld{sum\sub{enq}} fields of \var{root}.\fld{blocks} within that range to find $b_e$.}
Finally, the call to \op{GetEnqueue} \rev{at line \ref{findAnswer}} traces down through the tree to find the required \enqueue\ in a leaf.

>>>>

Having defined the fields required for tasks \ref{findinroot}, \ref{findrank} and \ref{findenqueue},
we can easily see how to construct a new block $B$ during a \op{Refresh} in $O(1)$ \rev{steps}.
A \op{Refresh} on \node\ $\var{v}$ reads the values $h_{\ell}$ and~$h_{r}$ of the \fld{head} fields of $\var{v}$'s children and stores 
$h_{\ell}-1$ and $h_{r}-1$ in $B.\eleft$ and $B.\eright$.
Then, we can compute 
\begin{eqnarray*}
B.\fld{sum\sub{enq}}&=&\var{v}.\fld{left}.\fld{blocks}[B.\eleft].\fld{sum\sub{enq}} \\
&&+ \var{v}.\fld{right}.\fld{blocks}[B.\eright].\fld{sum\sub{enq}}.
\end{eqnarray*}
For a block $B$ in the root, $B.\fld{size}$ is computed using the \fld{size} field of the previous block $B'$ and
the number of \enqueues\ and \dequeues\ in~$B$
\rev{as follows.  Before the operations of $B$, the queue has $B'.\fld{size}$ elements.
The \enqueues\ of $B$ add $B.\fld{sum\sub{enq}}-B'.\fld{sum\sub{enq}}$ elements.
Then the \dequeues\ of $B$ remove $B.\fld{sum\sub{deq}} - B'.\fld{sum\sub{deq}}$ elements.
(If subtracting this many elements would make the new size negative because there are more \dequeues\ than there are elements, then the new size is 0.)  Thus, we have}
\begin{eqnarray}
B.\fld{size} &=& \max(0, B'.\fld{size} + (B.\fld{sum\sub{enq}}-B'.\fld{sum\sub{enq}})\nonumber\\ 
			&& \hspace*{10mm}\mbox{ } - (B.\fld{sum\sub{deq}} - B'.\fld{sum\sub{deq}})).\label{size-eqn}
\end{eqnarray}

The only remaining field is $B.\fld{super}$.  When the block 
$B$ is created for a \node\ $\var{v}$, we do not yet know where its
superblock will eventually be installed in \var{v}'s parent.
So, we leave $B.\fld{super}$ blank.  
Soon after $B$ is installed,
%a call to \op{Advance} at line \ref{helpAdvance} or \ref{advance} by 
some process will
set \var{B}.\fld{super} to a value read from the \fld{head} field of \var{v}'s parent.
We shall show that this happens soon enough \rev{to ensure} that $B.\fld{super}$ can differ from the true index of \rev{$B$'s superblock}
by at most 1.
%After $B$ is installed 
%in $\var{v}.\fld{blocks}[\var{h}]$, processes cooperate to fill in $B.\fld{super}$ 
%when they attempt to advance $\var{v}.\fld{head}$ from $h$ to $h+1$,
%using a value they read from the \fld{head} field of $\var{v}$'s parent.
%As mentioned above, this might not be the exact index of $B$'s superblock, but we
%shall prove that it is close.

@@

\Function{void}{Enqueue}{\typ{Object} \var{e}} 
    \State \hangbox{let \var{B} be a new \typ{Block} with fields \fld{element} \assign\ \var{e},
		\fld{sum\sub{enq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{enq}}+1$,\\
		\fld{sum\sub{deq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{deq}}$}\label{enqNew}
    \State \Call{Append}{\var{B}}\Comment{\rev{append \op{Enqueue} to \var{leaf.}\fld{blocks} and propagate to root}}\label{enqAppend}
\EndFunction{Enqueue}

@@

\Function{Object}{Dequeue}{} 
    \State \hangbox{let \var{B} be a new \typ{Block} with fields \fld{element} \assign\ \nl,
	    \fld{sum\sub{enq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{enq}}$,\\
	    \fld{sum\sub{deq}} \assign\ $\var{leaf.}\fld{blocks}[\var{leaf.}\head-1].\fld{sum\sub{deq}}+1$}\label{deqNew}
    \State \Call{Append}{\var{B}}\Comment{\rev{append \dequeue\ to \var{leaf.}\fld{blocks} and propagate to root}}\label{deqAppend}
    \State $\langle \var{b}, \var{i}\rangle$ \assign\ \Call{IndexDequeue}{\var{leaf}, $\var{leaf.}\head-1$, $1$} \Comment{\rev{find where \op{Deqeueue} is in \var{root.}\fld{blocks}}}\label{invokeIndexDequeue}
    \State \Return{\Call{FindResponse}{\var{b, i}}}\Comment{\rev{find response \dequeue\ should return}}\label{deqRest}
\EndFunction{Dequeue}

@@

\Function{void}{Propagate}{\typ{Node} \var{v}} \Comment{propagate blocks from \var{v}'s children to root}
    \If{{\keywordfont not} \Call{Refresh}{\var{v}}} \label{firstRefresh}  \hfill \Comment{double refresh:  \rev{if first \op{Refresh} fails, try again}}
        \State \Call{Refresh}{\var{v}} \label{secondRefresh}
    \EndIf
    \If{$\var{v} \neq \var{root}$} \hfill \Comment{recurse up tree}
        \State \Call{Propagate}{\var{v}.\fld{parent}}
    \EndIf
\EndFunction{Propagate}

@@

\begin{figure*}
\begin{algorithmic}[1]
\setcounter{ALG@line}{39}
\Function{Block}{CreateBlock}{\typ{Node} \var{v}, \typ{int} \var{i}} \Comment{create new block $B$ for \op{Refresh} to install in \var{v}.\fld{blocks}[\var{i}]}
    \State \hangbox{let \var{B} be a new \typ{Block} with fields\label{initNewBlock}
    	\eleft\ \assign\ $\var{v}.\fld{left}.\head - 1$,\label{createEndLeft}
    	\eright\ \assign\ $\var{v}.\fld{right}.\head - 1$,\label{createEndRight}\\
		\fld{sum\sub{enq}} \assign\ \var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{enq}} + \var{v}.\fld{right.blocks}[\var{B}.\eright].\fld{sum\sub{enq}}\label{createSumEnq},\\
		\fld{sum\sub{deq}} \assign\ \var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{deq}} + \var{v}.\fld{right.blocks}[\var{B}.\eright].\fld{sum\sub{deq}}}\label{createSumDeq}
    \State \var{num\sub{enq}} \assign\ $\var{B}.\fld{sum\sub{enq}} - \var{v}.\fld{blocks}[\var{i}-1].\fld{sum\sub{enq}}$\label{computeNumEnq}
    \State \var{num\sub{deq}} \assign\ $\var{B}.\fld{sum\sub{deq}} - \var{v}.\fld{blocks}[\var{i}-1].\fld{sum\sub{deq}}$\label{computeNumDeq}
    \If{$\var{num\sub{enq}} + \var{num\sub{deq}} = 0$} \Return \nl\ \Comment{no blocks need to be propagated to \var{v}}\label{testEmpty} 
    \ElsIf{$\var{v} = \var{root}$} \var{B}.\fld{size} \assign\ max(0, $\var{v}.\fld{blocks}[\var{i}-1].\size\ + \var{num\sub{enq}} - \var{num\sub{deq}}$)\label{computeLength}
    \EndIf
    \State \Return \var{B}
\EndFunction{CreateBlock}

@@

    \If{\var{v} is a leaf node} \Return \var{v}.\fld{blocks}[\var{b}].\fld{element} \label{getBaseCase}
    \Else 
        \State \var{sum\sub{left}} \assign\ \var{v}.\fld{left.blocks}[\var{v}.\fld{blocks}[\var{b}].\eleft].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}$] from \var{v}.\fld{left}}
        \State \var{prev\sub{left}} \assign\ \var{v}.\fld{left.blocks}[\var{v}.\fld{blocks}[$\var{b}-1$].\eleft].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}-1$] from \var{v}.\fld{left}}
        \State \var{prev\sub{right}} \assign\ \var{v}.\fld{right.blocks}[\var{v}.\fld{blocks}[$\var{b}-1$].\eright].\fld{sum\sub{enq}} \Comment{\#\ \enqueues\ in \var{v}.\fld{blocks}[1..$\var{b}-1$] from \var{v}.\fld{right}}
        \If{$\var{i} \leq \var{sum\sub{left}} - \var{prev\sub{left}}$} \fld{dir} \assign\ \fld{left}\label{leftOrRight} \Comment{required \enqueue\ is in \var{v}.\fld{left}}
        \Else \Comment{required \enqueue\ is in \var{v}.\fld{right}}
            \State \fld{dir} \assign\ \fld{right}
            \State $\var{i}\ \assign\ \var{i} - (\var{sum\sub{left}} - \var{prev\sub{left}})$ \Comment{\rev{deduct \# \enqueues\ in $v.\fld{blocks}[b]$ that came from $v.\fld{left}$}}
        \EndIf \label{endChooseDir}
        \State \linecomment{find \enqueue's block in \var{v}.\fld{dir.blocks} and its rank within block}
        \State \hangbox{use binary search to find minimum $\var{b}'$ in range [\var{v}.\fld{blocks}[$\var{b}-1$].\edir+1..\var{v}.\fld{blocks}[\var{b}].\edir] such\\
        	 that $\var{v}.\fld{dir.blocks}[\var{b}'].\fld{sum\sub{enq}} \geq \var{i} + \var{prev\sub{dir}}$\label{getChild}}
        \State $\var{i}'$ \assign\ $\var{i} - (\var{v}.\fld{dir.blocks}[\var{b}'-1].\fld{sum\sub{enq}} - \var{prev\sub{dir}})$\label{getChildIndex}
        \State \Return\Call{GetEnqueue}{\var{v}.\fld{dir}, $\var{b}'$, $\var{i}'$} \label{getRecurse}
    \EndIf
\EndFunction{GetEnqueue}

@@

\opemph{CreateBlock}(\var{v, i}) is used
by \op{Refresh} to construct a new block $B$ to be installed in \var{v}.\fld{blocks}[\var{i}].
The \eleft\ and \eright\ fields store the indices of the last blocks appended to \var{v}'s
children, obtained by reading the \fld{head} index in \var{v}'s children.
Since the \fld{sum\sub{enq}} field should store the number of \enqueues\ in
\var{v}.\fld{blocks}[1..\var{i}] and these \enqueues\ come from \var{v}.\fld{left.blocks}[1..\var{B}.\eleft] and \var{v}.\fld{blocks}[1..\var{B}.\eright], line \ref{createSumEnq} sets
\fld{sum\sub{enq}} to the sum of $\var{v}.\fld{left.blocks}[\var{B}.\eleft].\fld{sum\sub{enq}}$ and $\var{v}.\fld{right}.\fld{blocks}[\var{B}.\eright].\fld{sum\sub{enq}}$.
Line \ref{computeNumEnq} sets \var{num\sub{enq}} to the number of \enqueues\ in the new block by
subtracting  the number of \enqueues\ in \var{v}.\fld{blocks}[$1..\var{i}-1$] from \var{B}.\fld{sum\sub{enq}}.
The values of \var{B}.\fld{sum\sub{deq}} and \var{num\sub{deq}} are computed similarly.
\rev{If the new block contains no operations, \op{CreateBlock} returns \nl\ at line \ref{testEmpty} to indicate
 there is no need to install~it.
Otherwise,} if \var{B}
is going to be installed in the root, line \ref{computeLength} \rev{uses Equation (\ref{size-eqn}) to compute $B$'s} \fld{size} field, which
represents the number of elements in the queue after the operations in the block are performed.

@@

To compute the response of the $i$th \dequeue\ in the $b$th block
of the root, \opemph{FindResponse}(\var{b, i}) determines at line \ref{checkEmpty} if the queue is empty.
If not, line \ref{computeE} computes the rank \var{e} of the
\op{Enqueue} whose argument is the \dequeue's response. 
\rev{The routine next searches for}
the index $b_e$ of the block in \var{root}.\fld{blocks} that contains 
the \var{e}th \enqueue.
Since the \enqueue\ is linearized before the \dequeue, \linebreak $b_e\leq b$.  To find the left end of the range \rev{to search for $b_e$, \op{FindResponse} does} a doubling search \cite{BY76}, comparing \var{e} to the \fld{sum\sub{enq}} fields at indices $b-1, b-2, b-4, b-8, \ldots$.
\rev{Then, the doubling search performs a binary search on the \fld{sum\sub{enq}} fields of \var{root}.\fld{blocks} within that range to find $b_e$.}
Finally, the call to \op{GetEnqueue} \rev{at line \ref{findAnswer}} traces down through the tree to find the required \enqueue\ in a leaf.
